---
title: "Happy Analysis"
author: "Harika Kovvuri"
date: "2023-02-18"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load New Package
```{r}
#install.packages("shiny")
#install.packages("shinydashboard")
```


#Load Programs
```{r}
library(tidyverse)
library(naniar)
library(janitor)
library(here)
library(shiny)
library(shinydashboard)
library(ggplot2)
```


#Load 2021 Data
```{r}
happy_2021 <-read.csv(here("happiness_data", "2021.csv")) %>% clean_names()
happy_2021
```

#Add 2021 Year Column
```{r}
happy_2021$year <- "2021"
happy_2021
```


#Load 2020 Data
```{r}
happy_2020 <-read.csv(here("happiness_data", "2020.csv")) %>% clean_names()
happy_2020
```

#Add 2020 Year Column
```{r}
happy_2020$year <-"2020"
happy_2020
```

#Retrieve Names
```{r}
names(happy_2020)
```
```{r}
names(happy_2021)
```

#Merge Data Sets
```{r}
merged_happy <-rbind(happy_2021,happy_2020)
```

```{r}
merged_happy
```

#Make New CSV File
```{r}
write.csv(merged_happy,"merged_happy.csv",row.names = F)
```

#Retrieve Newly Merged Data
```{r}
n_happy <- read.csv(here("happiness_data","merged_happy.csv"))
n_happy
```

#Structural Analysis of New Data

```{r}
names(n_happy)
```

```{r}
glimpse(n_happy)
```

#NA Summary
```{r}
naniar::miss_var_summary(n_happy)
```



# Create New Data Frame(Specified the Factors We Want to Analyze)
```{r}
narrow <- n_happy %>% 
  select(country_name,ladder_score,year,social_support,healthy_life_expectancy,freedom_to_make_life_choices,generosity,perceptions_of_corruption,logged_gdp_per_capita, dystopia_residual)
narrow
```

#2020 Narrow(Puts more focus on the stats from 2020)
```{r}
tw_narrow<-
  narrow%>% 
  filter(year=="2020")
tw_narrow
```

```{r}

names(tw_narrow)
```

#Create Stacked Bar Plot for 2020 Data

```{r}
names(tw_narrow)
```
#Reorder Dataframe involves pivoting longer for the 2020 data to create a bar plot.

```{r}
reorder <- tw_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
reorder
```

#Code for plot(this is specific to tw_narrow(2020 data))

```{r}
reorder %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2020", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
```

#Code for plot(this is specific to tw_narrow(2020 data))
```{r}
t1_narrow<-
  narrow%>% 
  filter(year=="2021")
t1_narrow
```


```{r}
revamp <- t1_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
revamp
```

```{r}
g2<-revamp %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
g2
```

#Slice Max and Min for Revamp(2021) and Reorder(2020)
##Revamp 2021 Max
```{r}
revamp5 <- t1_narrow %>% 
  top_n(5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score")

  revamp5 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
revamp5
```

