---
title: "Happy Analysis"
author: "Kyra Kong"
date: "2023-02-18"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load New Package
```{r}
#install.packages("shiny")
#install.packages("shinydashboard")
```


#Load Programs
```{r}
library(tidyverse)
library(naniar)
library(janitor)
library(here)
library(shiny)
library(shinydashboard)
library(ggplot2)
library(ggthemes)
```
```{r}
getwd()
```


#Load 2021 Data
```{r}
happy_2021 <-read.csv("/Users/merma/OneDrive/Desktop/happy/happiness_finder/happy_project/happiness_data/2021.csv") %>% clean_names()
happy_2021
```

#Add 2021 Year Column
```{r}
happy_2021$year <- "2021"
happy_2021
```


#Load 2020 Data
```{r}
happy_2020 <-read.csv("/users/merma/OneDrive/Desktop/happy/happiness_finder/happy_project/happiness_data/2020.csv") %>% clean_names()
happy_2020
```

#Add 2020 Year Column
```{r}
happy_2020$year <-"2020"
happy_2020
```

#Retrieve Names
```{r}
names(happy_2020)
```
```{r}
names(happy_2021)
```

#Merge Data Sets
```{r}
merged_happy <-rbind(happy_2021,happy_2020)
```

```{r}
merged_happy
```

#Make New CSV File
```{r}
write.csv(merged_happy,"merged_happy.csv",row.names = F)
```

#Retrieve Newly Merged Data
```{r}
n_happy <- read.csv("/users/merma/OneDrive/Desktop/happy/happiness_finder/happy_project/happiness_data/merged_happy.csv")
n_happy
```

#Structural Analysis of New Data

```{r}
names(n_happy)
```

```{r}
glimpse(n_happy)
```

#NA Summary
```{r}
naniar::miss_var_summary(n_happy)
```



# Create New Data Frame(Specified the Factors We Want to Analyze)
```{r}
narrow <- n_happy %>% 
  select(country_name,ladder_score,year,social_support,healthy_life_expectancy,freedom_to_make_life_choices,generosity,perceptions_of_corruption,logged_gdp_per_capita, dystopia_residual)
narrow
```

#2020 Narrow(Puts more focus on the stats from 2020)
```{r}
tw_narrow<-
  narrow%>% 
  filter(year=="2020")
tw_narrow
```

```{r}

names(tw_narrow)
```

#Create Stacked Bar Plot for 2020 Data

```{r}
names(tw_narrow)
```
#Reorder Dataframe involves pivoting longer for the 2020 data to create a bar plot.

```{r}
reorder <- tw_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
reorder
```

#Code for plot(this is specific to tw_narrow(2020 data))

```{r}
reorder %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2020", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
```

#Code for plot(this is specific to tw_narrow(2020 data))
```{r}
t1_narrow<-
  narrow%>% 
  filter(year=="2021")
t1_narrow
```


```{r}
revamp <- t1_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
revamp
```

```{r}
g2<-revamp %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
g2
```

#Slice Max and Min for Revamp(2021) and Reorder(2020)
##Revamp 2021 Max
```{r}
revamp5 <- t1_narrow %>% 
  top_n(5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score")

  revamp5 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
revamp5
```
##revamp 2021 min 
```{r}
revampbottom5 <- t1_narrow %>% 
  top_n(-5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score")
```

```{r}
 revampbottom5 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "Bottom 5 Countries Happiness 2021", x= "Countries", y="Factors")
revampbottom5
```
## reorder max 5
```{r}
reordermax5 <-tw_narrow %>% 
  top_n(5,ladder_score) %>% 
  group_by(country_name) %>% 
  pivot_longer(-c(country_name, year,ladder_score),
               names_to = "Factors",
               values_to= "Score")
```

```{r}
reordermax5 %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill= Factors))+
  geom_col(alpha =0.55)+
  coord_flip()+
  labs(title = "Happiness Score In 2020", x= "Countries", y= "Factors")
```

```{r}
reordermin <-tw_narrow %>% 
  top_n(-5,ladder_score) %>% 
  group_by(country_name) %>% 
  pivot_longer(-c(country_name, year,ladder_score),
               names_to = "Factors",
               values_to= "Score")
```

```{r}
reordermin %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill= Factors))+
  geom_col(alpha =0.55)+
  coord_flip()+
  labs(title= "Least Happiness Scores in 2020", x= "Countries", y= "Factors")
```

## analysis for bottom 5 countries ladder scores 
```{r}
tabyl(reordermin$country_name)
```

```{r}
tabyl(revampbottom5$country_name)
```
Similar country names are Afghanistan, Rwanda, and Zimbabwe.

Lesotho went from 11th least happiest country in 2020 to the 5th  least happiest country in 2021.
Botswana was in 7th least happiest country in 2020 and went to the 4th least happiest country in 2021.
Central African Republic was the 5th least happiest country in 2020 and was not listed in the dataset in 2021.


#ladder scores changes from 2020 to 2021

There's only one data set for Central African Republic

```{r}
merged_happy %>% 
  group_by(year) %>% 
  select(country_name, ladder_score, year) %>% 
  filter(country_name == "Central African Republic")
```
```{r}
reordermin %>% 
  ggplot(aes(country_name, y= ladder_score, fill= country_name))+
  geom_col()+
  coord_flip()
```
```{r}
revampbottom5 %>% 
  ggplot(aes(x=country_name, y= ladder_score, fill = country_name))+
  geom_col()+
  coord_flip()
```

```{r}
n_bottom_happy <-n_happy %>% 
  top_n(-30, ladder_score)
n_bottom_happy %>% 
  group_by(country_name)
```
Similar country names are Afghanistan, Rwanda, and Zimbabwe.

Lesotho went from 11th least happiest country in 2020 to the 5th  least happiest country in 2021.
Botswana was in 7th least happiest country in 2020 and went to the 4th least happiest country in 2021.
Central African Republic was the 5th least happiest country in 2020 and was not listed in the dataset in 2021.
South Sudan was 2nd most least happiest country in 2020 but not listed in the dataset in 2021.

```{r}
least_happy <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Afghanistan") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Afghanistan from 2020-2021")
least_happy
```
```{r}
second_least_happy <-n_bottom_happy %>% 
group_by(year) %>% 
  filter(country_name=="Zimbabwe") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Zimbabwe from 2020-2021")
second_least_happy
```

```{r}
third_least_happy <-n_bottom_happy %>% 
  group_by(year) %>% 
  filter(country_name=="Rwanda") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Rwanda from 2020-2021")
third_least_happy
```
```{r}
fourth_least_happy <-n_bottom_happy %>% 
  group_by(year) %>% 
  filter(country_name=="Botswana") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Botswana from 2020-2021")
fourth_least_happy
```
```{r}
fifth_least_happy <-n_bottom_happy %>% 
  group_by(year) %>% 
  filter(country_name=="Lesotho") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Lesotho from 2020-2021")
fifth_least_happy
```


# might make a faceted grid for these scores

