---
title: "Happy Analysis"
author: "Kyra Kong"
date: "2023-02-18"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load New Package
```{r}
#install.packages("shiny")
#install.packages("shinydashboard")
```


#Load Programs
```{r}
library(tidyverse)
library(naniar)
library(janitor)
library(here)
library(shiny)
library(shinydashboard)
library(ggplot2)
library(ggthemes)
```
```{r}
getwd()
```


#Load 2021 Data
```{r}
happy_2021 <-read.csv("/Users/merma/OneDrive/Desktop/happy/happiness_finder/happy_project/happiness_data/2021.csv") %>% clean_names()
happy_2021
```

#Add 2021 Year Column
```{r}
happy_2021$year <- "2021"
happy_2021
```


#Load 2020 Data
```{r}
happy_2020 <-read.csv("/users/merma/OneDrive/Desktop/happy/happiness_finder/happy_project/happiness_data/2020.csv") %>% clean_names()
happy_2020
```

#Add 2020 Year Column
```{r}
happy_2020$year <-"2020"
happy_2020
```

#Retrieve Names
```{r}
names(happy_2020)
```
```{r}
names(happy_2021)
```

#Merge Data Sets
```{r}
merged_happy <-rbind(happy_2021,happy_2020)
```

```{r}
merged_happy
```

#Make New CSV File
```{r}
write.csv(merged_happy,"merged_happy.csv",row.names = F)
```

#Retrieve Newly Merged Data
```{r}
n_happy <- read.csv("/users/merma/OneDrive/Desktop/happy/happiness_finder/happy_project/happiness_data/merged_happy.csv")
n_happy
```

#Structural Analysis of New Data

```{r}
names(n_happy)
```

```{r}
glimpse(n_happy)
```

#NA Summary
```{r}
naniar::miss_var_summary(n_happy)
```



# Create New Data Frame(Specified the Factors We Want to Analyze)
```{r}
narrow <- n_happy %>% 
  select(country_name,ladder_score,year,social_support,healthy_life_expectancy,freedom_to_make_life_choices,generosity,perceptions_of_corruption,logged_gdp_per_capita, dystopia_residual)
narrow
```

#2020 Narrow(Puts more focus on the stats from 2020)
```{r}
tw_narrow<-
  narrow%>% 
  filter(year=="2020")
tw_narrow
```

```{r}

names(tw_narrow)
```

#Create Stacked Bar Plot for 2020 Data

```{r}
names(tw_narrow)
```
#Reorder Dataframe involves pivoting longer for the 2020 data to create a bar plot.

```{r}
reorder <- tw_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
reorder
```

#Code for plot(this is specific to tw_narrow(2020 data))

```{r}
reorder %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2020", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
```

#Code for plot(this is specific to tw_narrow(2020 data))
```{r}
t1_narrow<-
  narrow%>% 
  filter(year=="2021")
t1_narrow
```


```{r}
revamp <- t1_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
revamp
```

```{r}
g2<-revamp %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
g2
```

#Slice Max and Min for Revamp(2021) and Reorder(2020)
##Revamp 2021 Max
```{r}
revamp5 <- t1_narrow %>% 
  top_n(5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score")

  revamp5 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
revamp5
```
##revamp 2021 min 
```{r}
revampbottom5 <- t1_narrow %>% 
  top_n(-5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score")
```

```{r}
 revampbottom5 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "Bottom 5 Countries Happiness 2021", x= "Countries", y="Factors")
revampbottom5
```
## reorder max 5
```{r}
reordermax5 <-tw_narrow %>% 
  top_n(5,ladder_score) %>% 
  group_by(country_name) %>% 
  pivot_longer(-c(country_name, year,ladder_score),
               names_to = "Factors",
               values_to= "Score")
```

```{r}
reordermax5 %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill= Factors))+
  geom_col(alpha =0.55)+
  coord_flip()+
  labs(title = "Happiness Score In 2020", x= "Countries", y= "Factors")
```

```{r}
reordermin <-tw_narrow %>% 
  top_n(-5,ladder_score) %>% 
  group_by(country_name) %>% 
  pivot_longer(-c(country_name, year,ladder_score),
               names_to = "Factors",
               values_to= "Score")
```

```{r}
reordermin %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill= Factors))+
  geom_col(alpha =0.55)+
  coord_flip()+
  labs(title= "Least Happiness Scores in 2020", x= "Countries", y= "Factors")+
  theme_tufte()
  
```

## analysis for bottom 5 countries ladder scores 
```{r}
tabyl(reordermin$country_name)
```

```{r}
tabyl(revampbottom5$country_name)
```
Similar country names are Afghanistan, Rwanda, and Zimbabwe.

Lesotho went from 11th least happiest country in 2020 to the 5th  least happiest country in 2021.
Botswana was in 7th least happiest country in 2020 and went to the 4th least happiest country in 2021.
Central African Republic was the 5th least happiest country in 2020 and was not listed in the dataset in 2021.


#ladder scores changes from 2020 to 2021

There's only one data set for Central African Republic

```{r}
merged_happy %>% 
  group_by(year) %>% 
  select(country_name, ladder_score, year) %>% 
  filter(country_name == "Central African Republic")
```
```{r}
reordermin %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill= country_name))+
  geom_col()+
  coord_flip()+
   theme_tufte()+
  labs(title= "Bottom Five Countries in 2020", y= "Ladder Score", x = "Country")
```
```{r}
revampbottom5 %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill = country_name))+
  geom_col()+
  coord_flip()+
  theme_tufte()+
  labs(title= "Bottom 5 Countries' Ladder Score in 2021", y= "Ladder Score", x= "Country Name")
```

```{r}
n_bottom_happy <-n_happy %>% 
  top_n(-30, ladder_score)
n_bottom_happy %>% 
  group_by(country_name)
```
Similar country names are Afghanistan, Rwanda, and Zimbabwe.

Lesotho went from 11th least happiest country in 2020 to the 5th least happiest country in 2021.
Botswana was in 7th least happiest country in 2020 and went to the 4th least happiest country in 2021.
Central African Republic was the 5th least happiest country in 2020 and was not listed in the dataset in 2021.
South Sudan was 2nd most least happiest country in 2020 but not listed in the dataset in 2021.

```{r}
least_happy <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Afghanistan") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Afghanistan from 2020-2021")
least_happy
```
```{r}
second_least_happy <-n_bottom_happy %>% 
group_by(year) %>% 
  filter(country_name=="Zimbabwe") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Zimbabwe from 2020-2021")
second_least_happy
```

```{r}
third_least_happy <-n_bottom_happy %>% 
  group_by(year) %>% 
  filter(country_name=="Rwanda") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Rwanda from 2020-2021")
third_least_happy
```
```{r}
fourth_least_happy <-n_bottom_happy %>% 
  group_by(year) %>% 
  filter(country_name=="Botswana") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Botswana from 2020-2021")
fourth_least_happy
```
```{r}
fifth_least_happy <-n_bottom_happy %>% 
  group_by(year) %>% 
  filter(country_name=="Lesotho") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Lesotho from 2020-2021")
fifth_least_happy
```



## app showing factors for bottom 5 countries

pivoting datasets to wider:
```{r}
bottom_5_2020 <- reordermin %>% 
  pivot_wider(names_from = "Factors",
              values_from= "Score")
```

```{r}
bottom5_2021 <-revampbottom5 %>% 
  pivot_wider(names_from = "Factors",
              values_from= "Score")
```


```{r}
bottom_5_2020 %>% 
  ggplot(aes(x= reorder(country_name,ladder_score),country_name, y= ladder_score, fill= country_name))+
  geom_col()+
  coord_flip()
```
# Social Support for 2020 and 2021

```{r}
names(n_bottom_happy)
```


```{r}

social_support_afghanistan <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Afghanistan") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Afghanistan from 2020-2021")
social_support_afghanistan

```


```{r}
social_support_zimbabwe <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Zimbabwe") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Zimbabwe from 2020-2021")
social_support_zimbabwe

```

```{r}
social_support_rwanda <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Rwanda") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Rwanda from 2020-2021")
social_support_rwanda

```

```{r}
social_support_botswana <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Botswana") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Botswana from 2020-2021")
social_support_botswana
```
```{r}
social_support_lesotho <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Lesotho") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Lesotho from 2020-2021")
social_support_lesotho
```
## Life Expectancy changes in bottom 5 countries from 2020-2021
```{r}
names(n_bottom_happy)
```


```{r}
life_expectancy_afghanistan <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Afghanistan") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Afghanistan from 2020-2021")
life_expectancy_afghanistan
```

```{r}
life_expectancy_zimbabwe <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Zimbabwe") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Zimbabwe from 2020-2021")
life_expectancy_zimbabwe
```
```{r}
life_expectancy_rwanda <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Rwanda") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Rwanda from 2020-2021")
life_expectancy_rwanda
```
```{r}
life_expectancy_botswana <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Botswana") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Botswana from 2020-2021")
life_expectancy_botswana
```

```{r}
life_expectancy_lesotho <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Lesotho") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Lesotho from 2020-2021")
life_expectancy_lesotho
```

## statistical analysis of the 3 factors selected: ladder score, life expectancy, and social support 

```{r}
n_bottom_happy_ladder_score <-n_bottom_happy %>% 
  select(country_name, ladder_score, year)

n_bottom_happy_ladder_score 
```
```{r}
n_bottom_happy_ladder_score<-n_bottom_happy_ladder_score %>% 
  pivot_wider(names_from= "year", names_prefix = "yr_", values_from = ladder_score) 

```
```{r}
n_bottom_happy_ladder_score <- n_bottom_happy_ladder_score %>% 
  select(country_name, yr_2020, yr_2021)
n_bottom_happy_ladder_score
```

```{r}
n_bottom_happy_ladder_score %>% 
  mutate(diff_ladd_score = (yr_2020-yr_2021)) %>% 
  arrange(desc(diff_ladd_score))
```


```{r}
n_bottom_happy_life_expectancy <-n_bottom_happy %>% 
  select(country_name, healthy_life_expectancy, year)

n_bottom_happy_life_expectancy
```
```{r}
n_bottom_happy_life_expectancy <-n_bottom_happy_life_expectancy  %>% 
  pivot_wider(names_from= year,names_prefix = "yr_",  values_from = healthy_life_expectancy) 
```
```{r}
n_bottom_happy_life_expectancy <- n_bottom_happy_life_expectancy %>% 
  select(country_name, yr_2020, yr_2021)
```

```{r}
n_bottom_happy_life_expectancy %>% 
  mutate(diff_life_exp = yr_2020-yr_2021) %>% 
  arrange(desc(diff_life_exp))
```
Out of the least happiest countries, the one with the most change in life expectancy is Sierra Leone, it decreased in life expectancy by about 0.78 of a year. Afghanistan had the least amount of change between 2021 and 2020, since the life expectancy barely increased by 0.097.

```{r}
n_bottom_social_support <-n_bottom_happy %>% 
  select(country_name, social_support, year)
n_bottom_social_support
```

```{r}
n_bottom_social_support <- n_bottom_social_support %>% 
  pivot_wider(names_from = year,names_prefix = "yr_",  values_from = social_support)
```

```{r}
n_bottom_social_support <- n_bottom_social_support %>% 
  select(country_name, yr_2020, yr_2021)
```


```{r}
n_bottom_social_support %>% 
  mutate(diff_soc_sup = yr_2020-yr_2021) %>% 
  arrange(desc(diff_soc_sup))
```
Out of the least unhappiest countries, Yemen had the most change in the
