---
title: "Happy Analysis"
author: "Harika Kovvuri"
date: "2023-02-18"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load New Package
```{r}
#install.packages("shiny")
#install.packages("ggthemes")
```


#Load Programs
```{r}
library(tidyverse)
library(naniar)
library(janitor)
library(here)
library(shiny)
library(shinydashboard)
library(ggplot2)
library(ggthemes)
```


#Load 2021 Data
```{r}
#happy_2021 <-read.csv(here("happiness_data", "2021.csv")) %>% clean_names()
#happy_2021
```

#Add 2021 Year Column
```{r}
#happy_2021$year <- "2021"
#happy_2021
```


#Load 2020 Data
```{r}
#happy_2020 <-read.csv(here("happiness_data", "2020.csv")) %>% clean_names()
#happy_2020
```

#Add 2020 Year Column
```{r}
#happy_2020$year <-"2020"
#happy_2020
```

#Retrieve Names
```{r}
#names(happy_2020)
```
```{r}
#names(happy_2021)
```

#Merge Data Sets
```{r}
#merged_happy <-rbind(happy_2021,happy_2020)
```

```{r}
#merged_happy
```

#Make New CSV File
```{r}
#write.csv(merged_happy,"merged_happy.csv",row.names = F)
```

#Retrieve Newly Merged Data
```{r}
n_happy <- read.csv(here("merged_happy.csv"))
n_happy
```

#Structural Analysis of New Data

```{r}
names(n_happy)
```

```{r}
glimpse(n_happy)
```

#NA Summary
```{r}
naniar::miss_var_summary(n_happy)
```



# Create New Data Frame(Specified the Factors We Want to Analyze)
```{r}
narrow <- n_happy %>% 
  select(country_name,ladder_score,year,social_support,healthy_life_expectancy,freedom_to_make_life_choices,generosity,perceptions_of_corruption,logged_gdp_per_capita, dystopia_residual)
narrow
```

#2020 Narrow(Puts more focus on the stats from 2020)
```{r}
tw_narrow<-
  narrow%>% 
  filter(year=="2020")
tw_narrow
```

```{r}

names(tw_narrow)
```

#Create Stacked Bar Plot for 2020 Data

```{r}
names(tw_narrow)
```
#Reorder Dataframe involves pivoting longer for the 2020 data to create a bar plot.

```{r}
reorder <- tw_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
reorder
```

#Code for plot(this is specific to tw_narrow(2020 data))

```{r}
reorder %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2020", x= "Countries", y="Factors")+
      theme_tufte()+
  theme(axis.text.y =element_text(size=3.5))
```

#Code for plot(this is specific to tw_narrow(2020 data))
```{r}
t1_narrow<-
  narrow%>% 
  filter(year=="2021")
t1_narrow
```


```{r}
revamp <- t1_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
revamp
```

```{r}
g2<-revamp %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
      theme_tufte()+
  theme(axis.text.y =element_text(size=3.5))
g2
```

#Slice Max and Min for Revamp(2021) and Reorder(2020)
##Revamp 2021 Max
```{r}
revamp5 <- t1_narrow %>% 
  top_n(5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score")

  revamp5 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
    theme_tufte()+
  theme(axis.text.y =element_text(size=3.5))
revamp5
```

##Revamp 2021 Min
```{r}
revamp6 <- t1_narrow %>% 
  top_n(-5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score")

  revamp6 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
    theme_tufte()+
  theme(axis.text.y =element_text(size=3.5))
revamp5
```

##Reorder 2020 Max
```{r}

reorder1 <- tw_narrow %>% 
   top_n(5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") 


  reorder1 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2020", x= "Countries", y="Factors")+
    theme_tufte()+
  theme(axis.text.y =element_text(size=3.5))
revamp5
```
##Reorder 2020 Min
```{r}
reorder2 <- tw_narrow %>% 
   top_n(-5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") 


  reorder2 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2020", x= "Countries", y="Factors")+
    theme_tufte()+
  theme(axis.text.y =element_text(size=3.5))
revamp5
```

#Statistical Analysis of the TOP 5 COUNTRIES
```{r}
#A point of interest is that in 2020 the fifth spot went to Norway but in 2021 it is replaced by the Netherlands. 

specific <-
  n_happy %>% 
  top_n(12,ladder_score) %>% 
  bind_rows(n_happy %>% 
              top_n(-12,ladder_score)) %>% filter(country_name != "Tanzania")
specific
  
```

```{r}
one<-specific %>% 
  group_by(year) %>% 
  filter(country_name=="Finland") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Finland from 2020-2021")
one
 
```

```{r}
two<-specific %>% 
  group_by(year) %>% 
  filter(country_name=="Denmark") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Denmark from 2020-2021")
two
```

```{r}
three<-specific %>% 
  group_by(year) %>% 
  filter(country_name=="Switzerland") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Switzerland from 2020-2021")
  three
```
```{r}
four<-specific %>% 
  group_by(year) %>% 
  filter(country_name=="Iceland") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Iceland from 2020-2021")
four
```
```{r}
five<-specific %>% 
  group_by(year) %>% 
  filter(country_name=="Netherlands") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Netherlands from 2020-2021")
five
```

```{r}
six<-specific %>% 
  group_by(year) %>% 
  filter(country_name=="Norway") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Norway from 2020-2021")
six
```

#Attempt an App
```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Changes in Ladder Score "),
  dashboardSidebar(),
  dashboardBody(
  fluidRow(
  box(title = "Happiest and Least Happiest Countries", width = 3,
  radioButtons("x", "Select Country", choices = c( "Finland", "Denmark", "Norway", "Netherlands", "Iceland","Switzerland","Zimbabwe","South Sudan","Rwanda","Central African Republic","Afghanistan","Lesotho","Botswana"), 
              selected = "Norway")

  
  ), # close the first box
  box(title = "Graph 1", width = 7,
  plotOutput("plot", width = "600px", height = "600px")
  ) # close the second box
  ) # close the row
  ) # close the dashboard body
) # close the ui

server <- function(input, output, session) { 
  
  output$plot <- renderPlot({
    specific %>% 
    filter(country_name==input$x) %>% 
    ggplot(aes(x=year,y=ladder_score)) +
      geom_line()+
      theme_tufte(base_size = 16)+
      theme(panel.background = element_rect(fill = "linen"))+
      ggtitle("Changes in Ladder Score by Year For the Happiest and Least Happiest Countries")+
       theme(plot.title = element_text(size = 16))+
      theme(axis.title.y = element_text(size = 15))+
      theme(axis.title.x = element_text(size = 15))+
      theme(axis.text.x = element_text(angle = 65, hjust = 1))+
      labs(x = "Year", y = "Ladder Score")
  })
}
  
 

shinyApp(ui, server)
```


