---
title: "Happy Analysis"
author: "Kyra Kong"
date: "2023-02-18"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load New Package
```{r}
#install.packages("shiny")
#install.packages("shinydashboard")
```


#Load Programs
```{r}
library(tidyverse)
library(naniar)
library(janitor)
library(here)
library(shiny)
library(shinydashboard)
library(ggplot2)
library(ggthemes)
```
```{r}
getwd()
```


#Load 2021 Data
```{r}
happy_2021 <-read.csv("../happiness_data/2021.csv") %>% clean_names()
happy_2021
```

#Add 2021 Year Column
```{r}
happy_2021$year <- "2021"
happy_2021
```


#Load 2020 Data
```{r}
happy_2020 <-read.csv("../happiness_data/2020.csv") %>% clean_names()
happy_2020
```

#Add 2020 Year Column
```{r}
happy_2020$year <-"2020"
happy_2020
```

#Retrieve Names
```{r}
names(happy_2020)
```
```{r}
names(happy_2021)
```

#Merge Data Sets
```{r}
merged_happy <-rbind(happy_2021,happy_2020)
```

```{r}
merged_happy
```

#Make New CSV File
```{r}
write.csv(merged_happy,"merged_happy.csv",row.names = F)
```

#Retrieve Newly Merged Data
```{r}
n_happy <- read.csv("merged_happy.csv")
n_happy
```

#Structural Analysis of New Data

```{r}
names(n_happy)
```

```{r}
glimpse(n_happy)
```

#NA Summary
```{r}
naniar::miss_var_summary(n_happy)
```



# Create New Data Frame(Specified the Factors We Want to Analyze)
```{r}
narrow <- n_happy %>% 
  select(country_name,ladder_score,year,social_support,healthy_life_expectancy,freedom_to_make_life_choices,generosity,perceptions_of_corruption,logged_gdp_per_capita, dystopia_residual)
narrow
```

#2020 Narrow(Puts more focus on the stats from 2020)
```{r}
tw_narrow<-
  narrow%>% 
  filter(year=="2020")
tw_narrow
```

```{r}

names(tw_narrow)
```

#Create Stacked Bar Plot for 2020 Data

```{r}
names(tw_narrow)
```
#Reorder Dataframe involves pivoting longer for the 2020 data to create a bar plot.

```{r}
reorder <- tw_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
reorder
```

#Code for plot(this is specific to tw_narrow(2020 data))

```{r}
plot2020 <-reorder %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2020", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))+
  theme_tufte(base_size=8, ticks = TRUE)



```
```{r}
plot2020
```

#Code for plot(this is specific to tw_narrow(2020 data))
```{r}
t1_narrow<-
  narrow%>% 
  filter(year=="2021")
t1_narrow
```


```{r}
revamp <- t1_narrow %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score") %>% 
  clean_names()
revamp
```

```{r}
plot2021 <-revamp %>% 
  group_by(country_name) %>% 
 ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=factors))+
   geom_col(alpha=0.55)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))+
  theme_tufte(base_size=8, ticks = TRUE)

```

#Slice Max and Min for Revamp(2021) and Reorder(2020)
##Revamp 2021 Max
```{r}
revamp5 <- t1_narrow %>% 
  top_n(5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score")

  revamp5 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "The Dependency of Happiness on 7 Factors in 2021", x= "Countries", y="Factors")+
  theme(axis.text.y =element_text(size=3.5))
revamp5
```
##revamp 2021 min 
```{r}
revampbottom5 <- t1_narrow %>% 
  top_n(-5,ladder_score) %>% 
  group_by(country_name) %>% 
pivot_longer (-c(country_name,year,ladder_score),
  names_to= "Factors",
  values_to="Score")
```

```{r}
 revampbottom5 %>% 
  ggplot(aes(x=reorder(country_name,ladder_score),country_name,y=ladder_score, fill=Factors))+
   geom_col(alpha=0.59)+
  coord_flip()+
  labs(title = "Bottom 5 Countries Happiness 2021", x= "Countries", y="Factors")
revampbottom5
```
## reorder max 5
```{r}
reordermax5 <-tw_narrow %>% 
  top_n(5,ladder_score) %>% 
  group_by(country_name) %>% 
  pivot_longer(-c(country_name, year,ladder_score),
               names_to = "Factors",
               values_to= "Score")
```

```{r}
reordermax5 %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill= Factors))+
  geom_col(alpha =0.55)+
  coord_flip()+
  labs(title = "Happiness Score In 2020", x= "Countries", y= "Factors")
```

```{r}
reordermin <-tw_narrow %>% 
  top_n(-5,ladder_score) %>% 
  group_by(country_name) %>% 
  pivot_longer(-c(country_name, year,ladder_score),
               names_to = "Factors",
               values_to= "Score")
```

```{r}
reordermin %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill= Factors))+
  geom_col(alpha =0.55)+
  coord_flip()+
  labs(title= "Least Happiness Scores in 2020", x= "Countries", y= "Factors")+
  theme_tufte()
  
```

## analysis for bottom 5 countries ladder scores 
```{r}
tabyl(reordermin$country_name)
```

```{r}
tabyl(revampbottom5$country_name)
```
Similar country names are Afghanistan, Rwanda, and Zimbabwe.

Lesotho went from 11th least happiest country in 2020 to the 5th  least happiest country in 2021.
Botswana was in 7th least happiest country in 2020 and went to the 4th least happiest country in 2021.
Central African Republic was the 5th least happiest country in 2020 and was not listed in the dataset in 2021.


#ladder scores changes from 2020 to 2021

There's only one data set for Central African Republic

```{r}
merged_happy %>% 
  group_by(year) %>% 
  select(country_name, ladder_score, year) %>% 
  filter(country_name == "Central African Republic")
```
```{r}
reordermin %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill= country_name))+
  geom_col()+
  coord_flip()+
   theme_tufte()+
  labs(title= "Bottom Five Countries in 2020", y= "Ladder Score", x = "Country")
```
```{r}
revampbottom5 %>% 
  ggplot(aes(x= reorder(country_name,ladder_score), country_name, y= ladder_score, fill = country_name))+
  geom_col()+
  coord_flip()+
  theme_tufte()+
  labs(title= "Bottom 5 Countries' Ladder Score in 2021", y= "Ladder Score", x= "Country Name")
```

```{r}
n_bottom_happy <-n_happy %>% 
  top_n(-30, ladder_score)
n_bottom_happy %>% 
  group_by(country_name)
```
Similar country names are Afghanistan, Rwanda, and Zimbabwe.

Lesotho went from 11th least happiest country in 2020 to the 5th least happiest country in 2021.
Botswana was in 7th least happiest country in 2020 and went to the 4th least happiest country in 2021.
Central African Republic was the 5th least happiest country in 2020 and was not listed in the dataset in 2021.
South Sudan was 2nd most least happiest country in 2020 but not listed in the dataset in 2021.

```{r}
least_happy <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Afghanistan") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Afghanistan from 2020-2021")
least_happy
```
```{r}
second_least_happy <-n_bottom_happy %>% 
group_by(year) %>% 
  filter(country_name=="Zimbabwe") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Zimbabwe from 2020-2021")
second_least_happy
```

```{r}
third_least_happy <-n_bottom_happy %>% 
  group_by(year) %>% 
  filter(country_name=="Rwanda") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Rwanda from 2020-2021")
third_least_happy
```
```{r}
fourth_least_happy <-n_bottom_happy %>% 
  group_by(year) %>% 
  filter(country_name=="Botswana") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Botswana from 2020-2021")
fourth_least_happy
```
```{r}
fifth_least_happy <-n_bottom_happy %>% 
  group_by(year) %>% 
  filter(country_name=="Lesotho") %>% 
  ggplot(aes(x=year,y=ladder_score))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Ladder Score",title="Trend in Ladder Scores for Lesotho from 2020-2021")
fifth_least_happy
```



## app showing factors for bottom 5 countries

pivoting datasets to wider:
```{r}
bottom_5_2020 <- reordermin %>% 
  pivot_wider(names_from = "Factors",
              values_from= "Score")
```

```{r}
bottom5_2021 <-revampbottom5 %>% 
  pivot_wider(names_from = "Factors",
              values_from= "Score")
```


```{r}
bottom_5_2020 %>% 
  ggplot(aes(x= reorder(country_name,ladder_score),country_name, y= ladder_score, fill= country_name))+
  geom_col()+
  coord_flip()
```
# Social Support for 2020 and 2021

```{r}
names(n_bottom_happy)
```


```{r}

social_support_afghanistan <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Afghanistan") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Afghanistan from 2020-2021")
social_support_afghanistan

```


```{r}
social_support_zimbabwe <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Zimbabwe") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Zimbabwe from 2020-2021")
social_support_zimbabwe

```

```{r}
social_support_rwanda <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Rwanda") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Rwanda from 2020-2021")
social_support_rwanda

```

```{r}
social_support_botswana <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Botswana") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Botswana from 2020-2021")
social_support_botswana
```
```{r}
social_support_lesotho <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Lesotho") %>% 
  ggplot(aes(x=year,y= social_support))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Social Support",title="Trend in Social Support for Lesotho from 2020-2021")
social_support_lesotho
```
## Life Expectancy changes in bottom 5 countries from 2020-2021
```{r}
names(n_bottom_happy)
```


```{r}
life_expectancy_afghanistan <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Afghanistan") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Afghanistan from 2020-2021")
life_expectancy_afghanistan
```

```{r}
life_expectancy_zimbabwe <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Zimbabwe") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Zimbabwe from 2020-2021")
life_expectancy_zimbabwe
```
```{r}
life_expectancy_rwanda <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Rwanda") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Rwanda from 2020-2021")
life_expectancy_rwanda
```
```{r}
life_expectancy_botswana <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Botswana") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Botswana from 2020-2021")
life_expectancy_botswana
```

```{r}
life_expectancy_lesotho <- n_bottom_happy %>% 
 group_by(year) %>% 
  filter(country_name=="Lesotho") %>% 
  ggplot(aes(x=year,y= healthy_life_expectancy))+
  geom_line()+
      theme_tufte()+
        theme(panel.background = element_rect(fill = "linen"))+
  labs(x = "Year", y = "Healthy Life Expectancy",title="Trend in Life Expectancy for Lesotho from 2020-2021")
life_expectancy_lesotho
```

## statistical analysis of the 3 factors selected: ladder score, life expectancy, and social support 

```{r}
n_bottom_happy_ladder_score <-n_bottom_happy %>% 
  select(country_name, ladder_score, year)

n_bottom_happy_ladder_score 
```
```{r}
n_bottom_happy_ladder_score<-n_bottom_happy_ladder_score %>% 
  pivot_wider(names_from= "year", names_prefix = "yr_", values_from = ladder_score) 

```

```{r}
n_bottom_happy_ladder_score <- n_bottom_happy_ladder_score %>% 
  select(country_name, yr_2020, yr_2021)
n_bottom_happy_ladder_score
```

```{r}
n_bottom_happy_ladder_score %>% 
  mutate(perc_change_ladder_score = (yr_2021-yr_2020)/(yr_2020)*100) %>% 
  arrange((perc_change_ladder_score))
```
```{r}
n_bottom_happy_ladder_score %>% 
  filter(country_name == "Afghanistan" | country_name == "South Sudan" | country_name == "Zimbabwe" | country_name == "Rwanda"| country_name == "Central African Republic" | country_name == "Botswana"| country_name == "Lesotho") %>%  
  mutate(perc_change_ladder_score = (yr_2021-yr_2020)/(yr_2020)*100) %>% 
  arrange((perc_change_ladder_score))
```


The one with the greatest decrease in ladder score from the most unhappiest countries from 2020 to 2021 is Zimbabwe (coming from 3rd place in 2020 to 2nd place in 2021). The one with the greatest increase in ladder score in the most unhappiest countries from 2020 to 2021 is Zambia with an 8.3% increase.

For the ones with the least happiest ladder scores:
Afghanistan decreased in ladder score by -1.71%
South Sudan had no data in 2021 so NA for percent change
Zimbabwe decreased in ladder score by -4.67%.
Rwanda increased in ladder score by 3.1%
Central African Republic had no data in 2021, so NA for percent change
Botswana decreased in ladder score by -0.34%.
Lesotho decreased in ladder score by -3.85%

```{r}
n_bottom_happy_life_expectancy <-n_bottom_happy %>% 
  select(country_name, healthy_life_expectancy, year)

n_bottom_happy_life_expectancy
```
```{r}
n_bottom_happy_life_expectancy <-n_bottom_happy_life_expectancy  %>% 
  pivot_wider(names_from= year,names_prefix = "yr_",  values_from = healthy_life_expectancy) 
```
```{r}
n_bottom_happy_life_expectancy <- n_bottom_happy_life_expectancy %>% 
  select(country_name, yr_2020, yr_2021)
```

```{r}
n_bottom_happy_life_expectancy %>% 
  filter(country_name == "Afghanistan" | country_name == "South Sudan" | country_name == "Zimbabwe" | country_name == "Rwanda"| country_name == "Central African Republic" | country_name == "Botswana"| country_name == "Lesotho")  %>% 
  mutate(perc_change_life_exp = (yr_2021-yr_2020)/(yr_2020)*100) %>% 
  arrange(perc_change_life_exp)
```
The country that decreased the most in healthy life expectancy is Afghanistan and the one that increased the most in healthy life expectancy is Sierra Leone. No information was found for Togo, Egypt, Central African Republic or South Sudan since they had no data in either one or the other datasets in 2020 or 2021.

For the ones with the least happiest countries' ladder scores:

Afghanistan decreased in healthy life expectancy by -1.84*10^-1%
South Sudan had no data in 2021, so NA for percent change for life expectancy
Zimbabwe increased in life expectancy by 1.049%
Rwanda increased in life expectancy by 4.92*10^-1%.
Central African Republic had no data in 2021, so NA for percent change in life expectancy
Botswana increased in life expectancy by 5.84x10^-1%
Lesotho increased in life expectancy by 1.45%

```{r}
n_bottom_social_support <-n_bottom_happy %>% 
  select(country_name, social_support, year)
n_bottom_social_support
```

```{r}
n_bottom_social_support <- n_bottom_social_support %>% 
  pivot_wider(names_from = year,names_prefix = "yr_",  values_from = social_support)
```

```{r}
n_bottom_social_support <- n_bottom_social_support %>% 
  select(country_name, yr_2020, yr_2021)
```


```{r}
n_bottom_social_support %>% 
  filter(country_name == "Afghanistan" | country_name == "South Sudan" | country_name == "Zimbabwe" | country_name == "Rwanda"| country_name == "Central African Republic" | country_name == "Botswana"| country_name == "Lesotho") %>% 
  mutate(percentchange_soc_sup = (yr_2021-yr_2020)/(yr_2020)*100) %>% 
  arrange(percentchange_soc_sup)
```
Out of the least unhappiest countries, the one that had the most decrease in social support was Haiti, decreased by -8.97% from 2020 to 2021 and the one that had the most increase in social support was Rwanda with 2.06%.

Out of the least happiest countries in 2020 and 2021:
Afghanistan:  decreased in social support by -1.56%
South Sudan: no info in 2021, NA for percent change in social support
Zimbabwe
Rwanda: increased in social support by 2.06%
Central African Republic: no info in 2021, NA for percent change in social support
Botswana: increased in social support by 0.62%
Lesotho: Increased in social support by 0.833%


## analysis for the top 5 countries

```{r}
specific <-
  n_happy %>% 
  top_n(12,ladder_score) %>% 
  bind_rows(n_happy %>% 
              top_n(-12,ladder_score)) %>% filter(country_name != "Tanzania")
specific
```

## ladder score 
Top 5 countries for 2020 and 2021:  Finland, Denmark, Switzerland, Iceland, Norway, Netherlands

```{r}
n_top_ladder <- specific %>% 
select(country_name, ladder_score, year)
n_top_ladder
```

```{r}
n_top_ladder <-n_top_ladder %>% 
  pivot_wider(names_from= "year", names_prefix = "yr_", values_from = ladder_score) 
```
```{r}
n_top_ladder <- n_top_ladder %>% 
  select(country_name, yr_2020, yr_2021)
```


```{r}
n_top_ladder %>% 
  filter(country_name == "Finland" | country_name == "Denmark" | country_name== "Switzerland" | country_name == "Iceland" | country_name == "Netherlands" | country_name == "Norway") %>% 
  mutate(perc_change_ladder_score = (yr_2021-yr_2020)/(yr_2020)*100) %>% 
  arrange(perc_change_ladder_score)
```


## social support

```{r}
n_top_soc_support <- specific %>% 
  select(country_name, social_support, year)
n_top_soc_support
```

```{r}
n_top_soc_support <- n_top_soc_support %>% 
   pivot_wider(names_from= "year", names_prefix = "yr_", values_from = social_support)
```
```{r}
n_top_soc_support <- n_top_soc_support %>% 
  select(country_name, yr_2020, yr_2021)
```
```{r}
n_top_soc_support %>% 
  filter(country_name == "Finland" | country_name == "Denmark" | country_name== "Switzerland" | country_name == "Iceland" | country_name == "Netherlands" | country_name == "Norway") %>% 
  mutate(perc_change_soc_supp = (yr_2021-yr_2020)/(yr_2020)*100) %>% 
  arrange(perc_change_soc_supp)
```


## healthy life expectancy

```{r}
n_top_life_exp <- specific %>% 
  select(country_name, healthy_life_expectancy, year)
n_top_life_exp
```

```{r}
n_top_life_exp <- n_top_life_exp %>% 
   pivot_wider(names_from= "year", names_prefix = "yr_", values_from = healthy_life_expectancy)
```
```{r}
n_top_life_exp <- n_top_life_exp %>% 
  select(country_name, yr_2020, yr_2021)
```
```{r}
n_top_life_exp %>% 
  filter(country_name == "Finland" | country_name == "Denmark" | country_name== "Switzerland" | country_name == "Iceland" | country_name == "Netherlands" | country_name == "Norway") %>% 
  mutate(perc_change_life_exp = (yr_2021-yr_2020)/(yr_2020)*100) %>% 
  arrange(perc_change_life_exp)
```